Regression Cube Analysis of Cohort Study Data

### 2015-03-30 draft.tex
crossed of ToDos on the whiteboard

Testing features for association with diseases using regression analysis is one of the most important epidemiological tools.
%%
%Using it to assess the statistical resilience of a hypothesis rarely involves \emph{more than three features} due to the required subject count.
Using it to assess the statistical resilience of a hypothesis rarely involves \emph{more than three features} due to the higher dimensional problem and the required subject count.
%%
% missing overview techniques ist zu hart; es gibt ja techniken (eine scatterplot matrix zb.)
Due to the amount of data and only limited overview techniques, possible correlations may be missed.
%%
Explorative analyses and overview visualizations of the data set as presented in prior work \cite{Klemm2014VIS} are not tailored to a specific target variable and mostly highlight correlations between variables, which are known to the domain expert (e.g., correlation between body size and spine shape).
%%
We incorporate the regression analysis, which is familiar to the domain experts, into overview visualizations to support a hypothesis-free analysis or an analysis towards a specific disease or hypothesis.
%%
This is achieved by providing template regression formulas, which are applied to all potential variable combinations.
%%
Since the notation is familiar to epidemiologists, they can rapidly include their domain knowledge into the analysis process.
%%
Difference views between regression formulas allow to assess the influences of individual variables on the process.

Our contributions are:
\begin{itemize}
	\item An overview visualization technique representing feature interactions using target features.
	\item Visualization techniques, which incorporate overview visualizations of all regression analyses at once as well as details on demand techniques for in-depth investigations of feature relationships.
	\item Freely adjustable regression formulas provide a simple, yet powerful way to adjust the regression analysis to specific hypotheses about the data.
	\item The analysis of confounding variables by providing comparison views between different formula results.
%	\item The open and web-based approach of the system allows for analysis of any data using the presented method.
	\item The expansion and adaption of our method to any data with our open web-based system.
\end{itemize}
%%
The term `Regression Cube' has been used for the detailed analysis of one or a small number of regression models \cite{Ahmadi, Chan}.
%%
They extend 2D scatter plots to display detailed information about a regression model.
%%
In behalf of that, we declare our visualization technique for comparing and analyzing a large number of different regression models with different input features as Regression Cube.
%%
\section{Epidemiological Background} \label{sec:Background}
%%
This section covers the epidemiological workflow and requirements.
%%
\subsection{Epidemiological Workflow} \label{EpidemiologicalWorkflow}
Epidemiological research is performed by experts from different academic disciplines, such as physicians, statisticians and medical computer scientists focusing on biometrics and image segmentation.
%%
Their goal is to derive disease-specific risk factors by assessing epidemiological features using statistical methods.
%%
As described by Thew et al. \cite{Thew2009} the epidemiological workflow is divided into these different steps:
\begin{itemize}
	\item Clinicians make observations in the daily practice, which are translated into hypotheses.
	\item Epidemiologists compile a list of variables depicting the hypothesis and include confounding variables.
	\item Statisticians assess the association of the derived features towards the investigated disease.
\end{itemize}
%%
Relative risks can be determined if a statistical resilient association of features towards a condition is extracted.
%%
They indicate the per-subject chance of developing the disease.
%%
Reproducibility of results is an epidemiological key requirement and guides all analyses steps.
%%
Statistical programs, such as \texttt{SPSS}, are used to analyze the data regarding the classical sequential epidemiological workflow. Hence, images are mostly used for the communication of results, rather than for providing insight into the data.
%%

In \cite{Klemm2014VIS} we described an Interactive Visual Analysis approach for image-centric cohort study data, which connects to the variable listing step.
%%
The methods aim to derive hypotheses through analysis of the data and observations about previously unknown feature correlations.
%%
Employing \emph{hypothesis generation} requires overview visualizations of feature correlations, which are not supported by standard statistical processors.
%%
In this work, we focus on a similar approach to derive insight and even new hypotheses through the data rather than only using it for a confirmatory analysis.
%An alternative data-driven approach is described as follows.
%%
% \paragraph{Data Driven Hypothesis-Generation-based Analysis.}
% We already described an Interactive Visual Analysis approach for image-centric cohort study data, which connects to the variable listing step \cite{Klemm2014VIS}.
% %%
% The methods aim to derive hypotheses through analysis of the data and observations about previously unknown feature correlations.
% %%
% Employing \emph{hypotheses generation} requires overview visualizations of feature correlations, which are not supported by standard statistical processors.

\subsection{Epidemiological Data} \label{sec:EpidemiologicalData}
%%
%\com{Bernhard: Due to overlaps to VIS'14 paper in this section, this can be shortened. E.g. "More discussion of data types and characteristics can be found in [...]"}
%%
There are many data acquisition modalities available to epidemiologists, which are suitable for different analysis types.
%%
%We focus on cohort studies are a major epidemiological tool for gathering epidemiological features.
We focus on cohort studies, which impose the largest data sets and yield a highly heterogenous and incomplete information space.
%%
Subject data are collected either towards a specific disease or with the widest range possible.
%%
The latter allows the data set to be assessed towards different diseases.
%%
The feature space comprises information about lifestyle, somatometric variables, medical parameters, genetic data as well as medical images.
%%
These features are derived through different modalities, such as questionnaires, medical examinations or laboratory analyses.
%%
Many features are sparse, such as follow-up questions about a medication or treatment of a certain disease.
%%
Other features are exclusive for a sub-group, such as women-specific questions, e.g., number of born children or period status.
%%
%\com{Restriction of amount of data because it must not be triangulated and because the ethics committees are very restrictive}
%%
%\paragraph{Data Types.}
%%
Medical status features or lifestyle factors are often of \emph{dichotomous} (binary) type.
%%
%The data space comprises continuous variables (somatometric variables, such as body weight) as well as categorical variables (e.g. graduation type).
%%
%The data heterogeneity has to be taken into account when the analysis method is chosen.
%%
Continuous data are often discretized (e.g. 10 year steps for age) to equalize the variable types and to simplify the method selection.
%%
However, this should be avoided, since it reduces the information space and also introduces a new information bias, as assumptions are modeled through the discretization.
%%

@@

%% Regression Cube: A Technique for Multidimensional Visual Exploration and Interactive Pattern Finding
We use the term Regression Cube for a comparative 3D representation of multiple regression models.
%% einheitlich 2D-XXX oder 2D XXX, meist ist es 2D XXX
Chan et al. \cite{Chan} use the same term to describe an extension of the 2D scatter plot representation of a linear regression model (incorporating solely metric features) to a 3D Cube.
%%
They group subjects using a set of interaction techniques as well as clustering algorithms to calculate sub-groups, which can then be compared using their cube representation.
%%
Similar to \cite{Piringer}, they focus on highlighting details of the included models rather than comparing models consisting of different features.
%%
Insight is derived by subject grouping, which spawns new cube correlations and therefore allows drilling down to the data.
%%
In contrast to this approach, we focus on comparing models using quality-of-fit measures and do not focus on subdividing the data sets to gain insight.
%%
We model expert knowledge through the definition of regression formulas, which is not the focus of Chan et al.
%%

@@

\paragraph{3D (Cube) View.}
%%
Introducing $Z$ creates a 3D \emph{Regression Cube} (Fig.~\ref{fig:Cube} b).
%%
$R^2$ values of each cube entry (\emph{voxel}) are mapped to opacity to reduce the overlap.
%%
The visualization of $R^2$ values derived from different regression cubes (e.g., $Z \sim X + Y$) is misleading, as they can be compared relatively, but not in precise numbers.
%%
% e.g. wenn man es wâ°hlen kann
% i.e. wenn es so ist
Therefore, the $R^2$ results of different regression methods are encoded using different colors (i.e., blue for linear regression and red for logistic regression).
%%
% Adverb immer vors Hauptverb oder Satzende
Thus, the cube can be easily extended using other regression types.
%%
For cubes with a fixed target feature, e.g., $Cancer \sim X + Y + Z$, no such encodings are required and the $z$ dimension can be compared directly.
%%
\\\\
Our goal is to create an overview visualization for a data set, but we also want to incorporate expert knowledge into the visualization by adapting the underlying formulas.
%%
These two approaches do not exclude each other, they rather underline the difference in purpose of the chosen formula.
%%
The different analysis approaches require different starting points using the \emph{Regression Cube}.

\subsection{Analysis Workflow} \label{sec:Workflow}
%%
\begin{figure}[htb]
 \centering
 \includegraphics[width=3.0in]{figures/workflow}
 \caption{
 Different workflow types of the analysis using \emph{Regression Cubes} (based on~\cite{Keim}).
 %%
 % The analysis starts with a set of features given by the data set.
 %%
 [\textbf{1.AF}] The workflow starts by declaring a formula to specify a hypothesis, or to use a predefined formula for a hypothesis-free analysis.
 %%
 [\textbf{2.SI}] The \emph{Regression Cube} is then visualized.
 %%
 The user has the option to either adjust the formula or to derive details on demand on a specific regression.
 %%
 [\textbf{3.ZF}] Insights into the data yield either an adjustment of the current formula or a selection of a difference view.
 %%
 The latter is used to compare regression cubes.
 %%
 [\textbf{4.DD}] Details about features using the 2D mosaic plot representation yield insights and hypotheses about feature relations.
 }
  \label{fig:Workflow}
\end{figure}
%%
%\com{Different Formulas work as different steps in the analysis (Fig.~\ref{fig:Workflow})}
%\com{Workflow for different analysis approaches.}
%\com{0. Select the Variables.}
%\com{1. Select the Formulas}
%\com{a) Analysis without hypothesis about the data set (z ~ x+y)}
%\com{b) Analysis without hypothesis with confounder (z ~ age)}
%\com{c) Analysis with hypothesis about the data set (Var ~ x+y+z)}
%\com{d) Use Difference View}
%\com{d) Clarify using confounder (Var ~ x)}
%\com{3. Find interesting pane.}
%\com{4. Get Details on Demand.}
%%
% â¹bergang ist recht holprig gewesen
Our Regression Cube model is well-suited for different workflow analysis techniques, based on the Visual Analytics (VA) Mantra of Keim et al. \cite{Keim}:

@@

\paragraph{Cube Comparison.}
%%
Regression Cubes can be compared by creating difference views.
%%
One cube (extracted with one formula) acts as reference.
%%
The absolute difference in $R^2$ values towards the second cube is calculated, yielding a difference cube showing only the differences between the two formulas.
%%
For example, it can be utilized for comparing the influence of a single feature towards the complete result (e.g., $Z \sim X + Y$ and $Z \sim X + Y + Income$)
%%
% \paragraph{Formula Types.}
% %%
% Input formulas reflect \emph{hypotheses} about the data.
% %%
% Using the operators, dynamic variables and dataset features, many different assumptions can be formulated.
% %%
% Here we show just a few examples on how formulas can serve different purposes.
%
% $Z \sim X + Y$ is the default formula.
% %%
% This cube represents all possible combinations of two independent features towards all features in the data set, since we do not know which feature(s) are of interest.
% %%
% Each slice represents a different target feature.
% %%
% It is therefore suitable for a \emph{hypothesis-free} explorative analysis to give a general impression about relationships in the data set.

@@

\paragraph{Formula Editor [1.AF, 3.ZF].}
%%
After uploading the data, the user can specify a formula or use the default ($Z \sim X + Y$).
%%
Entering a formula is provided via text input.
%%
On formula input, a context modal displays all data set variables as well as the available operators and their function.
%%
This allows to comprehend the function of the underlying formula for users without statistical background about regression analysis and its notation.
%%
Auto-completing input variables also simplifies the approach and works as spell check of variable names.
%%

@@

Therefore, \emph{UN} started the \textbf{1.AF} step using the dichotomized MRI fat liver concentration using the formula $mrt\_liverfat\_s2 \sim X + Y + Z$ for \emph{male} subjects.

@@

The \textbf{2.SI} step using the 3D cube representations locates hotspots at the end of the cube (Fig.~\ref{fig:Application} a).

@@

The Zoom, Filter and Analyze Further Step (\textbf{3.ZF}) was realized by slicing through the cube using the mouse input to inspect the hot spots.

@@

Analyzing the mosaic plot (\textbf{4.DD}) revealed high correlations for somatometric features, hepatic steatosis indicator-features as well as laboratory values, such as creatinine (used renal retention parameter) and uric acid (among others used as gout and diabetes risk factors) magnitudes.

@@

\textbf{Analysis of non-discretized target feature.}
%%
%Since our method can analyze numerical target features, as opposed to decision trees used by \emph{UN} before, the analysis was conducted again for the not-dichotomized target using the same formula.
Since our method can assess numerical target features, the analysis was conducted again for the not-dichotomized target using the same formula.
%%
The 3D cube representation already showed lower $R^2$ values in general. However, the analysis is now based on linear regression and the $R^2$ values cannot be directly compared.
%%
The correlation hot spots matched with the ones from the dichotomous target, but were generally lower ($R^2$ of $0.37$ for somatometric features as opposed to $0.58$).
%%
We assume that the bias introduced by dichotomizing the fat liver content enforces the findings towards liver diseases, while using the numerical features is less expressive.
%%
\textbf{Interleukin-6 correlation with liver fat.}
%%
During the analysis, one hot spot was always observable in the \emph{SI} and \emph{ZF} incorporating a high \emph{Interleukin-6} (\emph{IL-6}) correlation with liver fat values (Fig.~\ref{fig:Application} b).
%%
The correlation was high for both the dichotomized and continuous target feature.
%%
The literature described relations between \emph{IL-6} and liver cancer \cite{He2013} as well as chronic liver diseases \cite{Streetz2003}.
%%
For mice, strong effects of \emph{IL-6} with hepatic steatosis were described \cite{Hong2004}.
%%
The finding is subject of further analysis.
%%
\subsection{Case 2: Hypothesis-free Analysis of the Breast Cancer Data Set}
%%
The analysis aims to find relationships on the breast fat data using features derived from the mammography analysis.
%%
Relationships between the share of parenchyma tissue on the overall breast volume are of high interest \cite{Mccormack2006}.
%%

The \textbf{1.AF} was started by \emph{KH} using the default formula for hypothesis-free analysis ($Z \sim X + Y$).
%%
At first, she was interested in correlations with the \emph{parenchyma tissue} percentage, which was selected through the drop-down for the z-axis (\textbf{2.SI}).
%%
She observed strong correlations with \emph{age}, \emph{body fat percentage}, \emph{hip} and \emph{waist circumference} as well as \emph{menstrual period} or \emph{pregnancy status} as expected (Fig.~\ref{fig:Application} right).
%%
Women with higher \emph{body fat} also have a larger \emph{breast fat percentage}, which also correlates with other somatometric features.
%%
\emph{Age} is a strong influencing factor, as breast tissue and subsequently the parenchyma tissue degrades over time.
%%
\emph{KH} proceeded using \textbf{2.SI} and \textbf{3.DD} to check for relationships for different target features, such as current \emph{hormone replacement therapy}, \emph{BI-RADS} (classification of the mammography findings) as well as different diseases, such as \emph{diabetes} or \emph{gout}.
%%
She observed relationships matching her expectations and expert knowledge.
%%
One unexpected relationship was observed between \emph{breast lesions} and \emph{menstruation cycle} towards \emph{spiral contraception} ($R^2$ of $0.77$).
%%
\emph{KH} proceeded with a detailed analysis of the parenchyma tissue.
%\emph{KH} derived at the hypotheses: \emph{Somatometric features impose the major influences on the parenchyma tissue.}
%%

\textbf{Detailed Breast Parenchyma Analysis.}
%%
The analysis was conducted by calculating the cube $Parenchyma\_Percentage \sim X + Y + Z$.
%%
Using the 3D cube, \emph{KH} observed several hotspots.
%%
Navigating to them using the slicing facility of the 3D visualization highlighted features of high influence, such as image-derived features, as \emph{glandular tissue density} and \emph{parenchyma segmentation} metrics.
%%
Also, strong correlations were observed in the \emph{diabetes} slice, confirming expectations of \emph{KH} w.r.t. to its strong influence to the parenchyma tissue.
%%
A surprising find were the strong correlations with \emph{kidney disorder} ($R^2$ values around $0.9$).
%%
The \textbf{3.DD} analysis however showed only 8 subjects with \emph{kidney disorder}.
%%
Too few subjects impose the risk of a biased finding.
%%
The correlation was noted and will be investigated further in the near future using a extensive data set.
%%
Lastly \emph{KH} assessed the influence of contraception related features, such as use \emph{birth control pills} or the \emph{spiral}, but found no significant correlations with the parenchyma tissue.
%%
Other consumption behavior features, such as \emph{alcohol intake} also yield no elevated $R^2$ values.
%%
\emph{KH} remarked that these features are believed to have an impact on the parenchyma tissue, but they are less reliability since they are self-reported.
%%
%mamma anamnesis breast fed, contraception (spirale, pille), alcohol intake
%Lastly, \emph{KH} assessed the influence of image-derived features on the parenchyma tissue.
%%
%She observed expected results 
%Using the slicing mode to fast 
%Hypotheses
%Detailed analysis: Cholesterin, influence of image-derived feautres, classification of the BI-Rads (how well did the manual segmentation perform)
\subsection{Further Feedback and Lessons Learned} \label{Lessons Learned}
%%
Our method was well received among the domain experts.
%%
For the first time, they were able to derive an overview visualization custom tailored to underlying assumptions.
%%
\emph{KH} noted the ease of use, which "\emph{converts data sets into a feasible form}".
%%
She highlighted the efficiency of combining fast target feature selection with visually highlighting interesting results, enabling rapid analysis cycles.
%%
To get nearly similar results she would had to spend hours using \texttt{SPSS} and also might have missed many interesting hot spots during this process.
%She noted that she usually would have to spend hours using \texttt{SPSS} to assess the data in the detail we presented.
%%
\emph{TI} highlighted the ability to simultaneously analyze thousands of regression models while maintaining little time expenses for rating them.
%The novel way of modeling knowledge to derive hypothesis driven overview visualizations allow for detailed analysis of target features, either disease indicators or even image-derived features.
%%
%TI commented on the

>>>>

Testing features for association with diseases using regression analysis is one of the most important epidemiological tools.
%%
%Using it to assess the statistical resilience of a hypothesis rarely involves \emph{more than three features} due to the required subject count.
Using it to assess the statistical resilience of a hypothesis rarely involves \emph{more than three features} due to the higher dimensional problem and the required subject count.
%%
% missing overview techniques ist zu hart; es gibt ja techniken (eine scatterplot matrix zb.)
Due to the amount of data and only limited overview techniques, possible correlations may be missed.
%%
Explorative analyses and overview visualizations of the data set as presented in prior work \cite{Klemm2014VIS} are not tailored to a specific target feature and mostly highlight correlations between features, which are known to the domain expert (e.g., correlation between body size and spine shape).
%%
We incorporate the regression analysis, which is familiar to the domain experts, into overview visualizations to support a hypothesis-free analysis or an analysis towards a specific disease or hypothesis.
%%
This is achieved by providing template regression formulas, which are applied to all potential feature combinations.
%%
Since the notation is familiar to epidemiologists, they can rapidly include their domain knowledge into the analysis process.
%%
Difference views between regression formulas allow to assess the influences of individual features on the process.

Our contributions are:
\begin{itemize}
	\item An overview visualization technique representing feature interactions using target features.
	\item Visualization techniques, which incorporate overview visualizations of all regression analyses at once as well as details on demand techniques for in-depth investigations of feature relationships.
	\item Freely adjustable regression formulas provide a simple, yet powerful way to adjust the regression analysis to specific hypotheses about the data.
	\item The analysis of confounding features by providing comparison views between different formula results.
%	\item The open and web-based approach of the system allows for analysis of any data using the presented method.
	\item The expansion and adaption of our method to any data with our open web-based system.
\end{itemize}
%%
The term `Regression Cube' has been used for the detailed analysis of one or a small number of regression models \cite{Ahmadi, Chan}.
%%
They extend 2D scatter plots to display detailed information about a regression model.
%%
In behalf of that, we declare our visualization technique for comparing and analyzing a large number of different regression models with different input features as \emph{Regression Cube}.
%%
\section{Epidemiological Background} \label{sec:Background}
%%
This section covers the epidemiological workflow and requirements.
%%
\subsection{Epidemiological Workflow} \label{EpidemiologicalWorkflow}
Epidemiological research is performed by experts from different academic disciplines, such as physicians, statisticians and medical computer scientists focusing on biometrics and image segmentation.
%%
Their goal is to derive disease-specific risk factors by assessing epidemiological features using statistical methods.
%%
As described by Thew et al. \cite{Thew2009} the epidemiological workflow is divided into these different steps:
\begin{itemize}
	\item Clinicians make observations in the daily practice, which are translated into hypotheses.
	\item Epidemiologists compile a list of features depicting the hypothesis and include confounding features.
	\item Statisticians assess the association of the derived features towards the investigated disease.
\end{itemize}
%%
Relative risks can be determined if a statistical resilient association of features towards a condition is extracted.
%%
They indicate the per-subject chance of developing the disease.
%%
Reproducibility of results is an epidemiological key requirement and guides all analyses steps.
%%
Statistical programs, such as \texttt{SPSS}, are used to analyze the data regarding the classical sequential epidemiological workflow. Hence, images are mostly used for the communication of results, rather than for providing insight into the data.
%%

In \cite{Klemm2014VIS} we described an Interactive Visual Analysis approach for image-centric cohort study data, which connects to the feature listing step.
%%
The methods aim to derive hypotheses through analysis of the data and observations about previously unknown feature correlations.
%%
Employing \emph{hypothesis generation} requires overview visualizations of feature correlations, which are not supported by standard statistical processors.
%%
In this work, we focus on a similar approach to derive insight and even new hypotheses through the data rather than only using it for a confirmatory analysis.
%An alternative data-driven approach is described as follows.
%%
% \paragraph{Data Driven Hypothesis-Generation-based Analysis.}
% We already described an Interactive Visual Analysis approach for image-centric cohort study data, which connects to the variable listing step \cite{Klemm2014VIS}.
% %%
% The methods aim to derive hypotheses through analysis of the data and observations about previously unknown feature correlations.
% %%
% Employing \emph{hypotheses generation} requires overview visualizations of feature correlations, which are not supported by standard statistical processors.

\subsection{Epidemiological Data} \label{sec:EpidemiologicalData}
%%
%\com{Bernhard: Due to overlaps to VIS'14 paper in this section, this can be shortened. E.g. "More discussion of data types and characteristics can be found in [...]"}
%%
There are many data acquisition modalities available to epidemiologists, which are suitable for different analysis types.
%%
%We focus on cohort studies are a major epidemiological tool for gathering epidemiological features.
We focus on cohort studies, which impose the largest data sets and yield a highly heterogenous and incomplete information space.
%%
Subject data are collected either towards a specific disease or with the widest range possible.
%%
The latter allows the data set to be assessed towards different diseases.
%%
The feature space comprises information about lifestyle, somatometric features, medical parameters, genetic data as well as medical images.
%%
These features are derived through different modalities, such as questionnaires, medical examinations or laboratory analyses.
%%
Many features are sparse, such as follow-up questions about a medication or treatment of a certain disease.
%%
Other features are exclusive for a sub-group, such as women-specific questions, e.g., number of born children or period status.
%%
%\com{Restriction of amount of data because it must not be triangulated and because the ethics committees are very restrictive}
%%
%\paragraph{Data Types.}
%%
Medical status features or lifestyle factors are often of \emph{dichotomous} (binary) type.
%%
%The data space comprises continuous variables (somatometric variables, such as body weight) as well as categorical variables (e.g. graduation type).
%%
%The data heterogeneity has to be taken into account when the analysis method is chosen.
%%
Continuous data are often discretized (e.g. 10 year steps for age) to equalize the feature types and to simplify the method selection.
%%
However, this should be avoided, since it reduces the information space and also introduces a new information bias, as assumptions are modeled through the discretization.
%%

@@

%% Regression Cube: A Technique for Multidimensional Visual Exploration and Interactive Pattern Finding
We use the term `Regression Cube' for a comparative 3D representation of multiple regression models.
%% einheitlich 2D-XXX oder 2D XXX, meist ist es 2D XXX
Chan et al. \cite{Chan} use the same term to describe an extension of the 2D scatter plot representation of a linear regression model (incorporating solely metric features) to a 3D Cube.
%%
They group subjects using a set of interaction techniques as well as clustering algorithms to calculate sub-groups, which can then be compared using their cube representation.
%%
Similar to \cite{Piringer}, they focus on highlighting details of the included models rather than comparing models consisting of different features.
%%
Insight is derived by subject grouping, which spawns new cube correlations and therefore allows drilling down to the data.
%%
In contrast to this approach, we focus on comparing models using quality-of-fit measures and do not focus on subdividing the data sets to gain insight.
%%
We model expert knowledge through the definition of regression formulas, which is not the focus of Chan et al.
%%

@@

\paragraph{3D (Cube) View.}
%%
Introducing $Z$ creates a 3D \emph{Regression Cube} (Fig.~\ref{fig:Cube} b).
%%
$R^2$ values of each cube entry (\emph{voxel}) are mapped to opacity to reduce the overlap.
%%
The visualization of $R^2$ values derived from different regression cubes (e.g., $Z \sim X + Y$) is misleading, as they can be compared relatively, but not in precise numbers.
%%
% e.g. wenn man es wâ°hlen kann
% i.e. wenn es so ist
Therefore, the $R^2$ results of different regression methods are encoded using different colors (i.e., blue for linear regression and red for logistic regression).
%%
% Adverb immer vors Hauptverb oder Satzende
Thus, the cube can be easily extended using other regression types.
%%
For cubes with a fixed target feature, e.g., $Cancer \sim X + Y + Z$, no such encodings are required and the $z$ dimension can be compared directly.
%%
The color for both 2D and 3D view encodes the type of the regression analysis (blue for logistic regressions, orange for linear regressions).
%%
\\\\
%%
Our goal is to create an overview visualization for a data set, but we also want to incorporate expert knowledge into the visualization by adapting the underlying formulas.
%%
These two approaches do not exclude each other, they rather underline the difference in purpose of the chosen formula.
%%
The different analysis approaches require different starting points using the \emph{Regression Cube}.

\subsection{Analysis Workflow} \label{sec:Workflow}
%%
\begin{figure}[htb]
 \centering
 \includegraphics[width=3.0in]{figures/workflow}
 \caption{
 Different workflow types of the analysis using \emph{Regression Cubes} (based on~\cite{Keim}).
 %%
 % The analysis starts with a set of features given by the data set.
 %%
 [\textbf{1.AF}] The workflow starts by declaring a formula to specify a hypothesis, or to use a predefined formula for a hypothesis-free analysis.
 %%
 [\textbf{2.SI}] The \emph{Regression Cube} is then visualized.
 %%
 The user has the option to either adjust the formula or to derive details on demand on a specific regression.
 %%
 [\textbf{3.ZF}] Insights into the data yield either an adjustment of the current formula or a selection of a difference view.
 %%
 The latter is used to compare regression cubes.
 %%
 [\textbf{4.DD}] Details about features using the 2D mosaic plot representation yield insights and hypotheses about feature relations.
 }
  \label{fig:Workflow}
\end{figure}
%%
%\com{Different Formulas work as different steps in the analysis (Fig.~\ref{fig:Workflow})}
%\com{Workflow for different analysis approaches.}
%\com{0. Select the Variables.}
%\com{1. Select the Formulas}
%\com{a) Analysis without hypothesis about the data set (z ~ x+y)}
%\com{b) Analysis without hypothesis with confounder (z ~ age)}
%\com{c) Analysis with hypothesis about the data set (Var ~ x+y+z)}
%\com{d) Use Difference View}
%\com{d) Clarify using confounder (Var ~ x)}
%\com{3. Find interesting pane.}
%\com{4. Get Details on Demand.}
%%
% â¹bergang ist recht holprig gewesen
Our \emph{Regression Cube} model is well-suited for different workflow analysis techniques, based on the Visual Analytics (VA) Mantra of Keim et al. \cite{Keim}:

@@

\paragraph{Cube Comparison.}
%%
\emph{Regression Cubes} can be compared by creating difference views.
%%
One cube (extracted with one formula) acts as reference.
%%
The absolute difference in $R^2$ values towards the second cube is calculated, yielding a difference cube showing only the differences between the two formulas.
%%
For example, it can be utilized for comparing the influence of a single feature towards the complete result (e.g., $Z \sim X + Y$ and $Z \sim X + Y + Income$)
%%
% \paragraph{Formula Types.}
% %%
% Input formulas reflect \emph{hypotheses} about the data.
% %%
% Using the operators, dynamic variables and dataset features, many different assumptions can be formulated.
% %%
% Here we show just a few examples on how formulas can serve different purposes.
%
% $Z \sim X + Y$ is the default formula.
% %%
% This cube represents all possible combinations of two independent features towards all features in the data set, since we do not know which feature(s) are of interest.
% %%
% Each slice represents a different target feature.
% %%
% It is therefore suitable for a \emph{hypothesis-free} explorative analysis to give a general impression about relationships in the data set.

@@

\paragraph{Formula Editor [1.AF, 3.ZF].}
%%
After uploading the data, the user can specify a formula or use the default ($Z \sim X + Y$).
%%
Entering a formula is provided via text input.
%%
On formula input, a context modal displays all data set features as well as the available operators and their function.
%%
This allows to comprehend the function of the underlying formula for users without statistical background about regression analysis and its notation.
%%
Auto-completing input features also simplifies the approach and works as spell check of feature names.
%%

@@

Therefore, \emph{UN} started the [\textbf{1.AF}] step using the dichotomized MRI fat liver concentration using the formula $mrt\_liverfat\_s2 \sim X + Y + Z$ for \emph{male} subjects.

@@

The [\textbf{2.SI}] step using the 3D cube representations locates hotspots at the end of the cube (Fig.~\ref{fig:Application} a).

@@

The Zoom, Filter and Analyze Further Step [\textbf{3.ZF}] was realized by slicing through the cube using the mouse input to inspect the hot spots.

@@

Analyzing the mosaic plot [\textbf{4.DD}] revealed high correlations for somatometric features, hepatic steatosis indicator-features as well as laboratory values, such as creatinine (used renal retention parameter) and uric acid (among others used as gout and diabetes risk factors) magnitudes.

@@

\textbf{Analysis of non-discretized target feature.}
%%
%Since our method can analyze numerical target features, as opposed to decision trees used by \emph{UN} before, the analysis was conducted again for the not-dichotomized target using the same formula.
Since our method can assess numerical target features, the analysis was conducted again for the not-dichotomized target using the same formula.
%%
The 3D cube representation already showed lower $R^2$ values in general. However, the analysis is now based on linear regression and the $R^2$ values cannot be directly compared.
%%
The correlation hot spots matched with the ones from the dichotomous target, but were generally lower ($R^2$ of $0.37$ for somatometric features as opposed to $0.58$).
%%
We assume that the bias introduced by dichotomizing the fat liver content enforces the findings towards liver diseases, while using the numerical features is less expressive.

\textbf{Interleukin-6 correlation with liver fat.}
%%
During the analysis, one hot spot was always observable in the [\textbf{2.SI}] and [\textbf{3.ZF}] incorporating a high \emph{Interleukin-6} (\emph{IL-6}) correlation with liver fat values (Fig.~\ref{fig:Application} b).
%%
The correlation was high for both the dichotomized and continuous target feature.
%%
The literature described relations between \emph{IL-6} and liver cancer \cite{He2013} as well as chronic liver diseases \cite{Streetz2003}.
%%
For mice, strong effects of \emph{IL-6} with hepatic steatosis were described \cite{Hong2004}.
%%
The finding is subject of further analysis.
%%
\subsection{Case 2: Hypothesis-free Analysis of the Breast Cancer Data Set}
%%
The analysis aims to find relationships on the breast fat data using features derived from the mammography analysis.
%%
Relationships between the share of parenchyma tissue on the overall breast volume are of high interest \cite{Mccormack2006}.
%%

The [\textbf{1.AF}] was started by \emph{KH} using the default formula for hypothesis-free analysis ($Z \sim X + Y$).
%%
At first, she was interested in correlations with the \emph{parenchyma tissue} percentage, which was selected through the drop-down for the z-axis [\textbf{2.SI}].
%%
She observed strong correlations with \emph{age}, \emph{body fat percentage}, \emph{hip} and \emph{waist circumference} as well as \emph{menstrual period} or \emph{pregnancy status} as expected (Fig.~\ref{fig:Application} right).
%%
Women with higher \emph{body fat} also have a larger \emph{breast fat percentage}, which also correlates with other somatometric features.
%%
\emph{Age} is a strong influencing factor, as breast tissue and subsequently the parenchyma tissue degrades over time.
%%
\emph{KH} proceeded using [\textbf{3.ZF}] and [\textbf{4.DD}] to check for relationships for different target features, such as current \emph{hormone replacement therapy}, \emph{BI-RADS} (classification of the mammography findings) as well as different diseases, such as \emph{diabetes} or \emph{gout}.
%%
She observed relationships matching her expectations and expert knowledge.
%%
One unexpected relationship was observed between \emph{breast lesions} and \emph{menstruation cycle} towards \emph{spiral contraception} ($R^2$ of $0.77$).
%%
\emph{KH} proceeded with a detailed analysis of the parenchyma tissue.
%\emph{KH} derived at the hypotheses: \emph{Somatometric features impose the major influences on the parenchyma tissue.}
%%

\textbf{Detailed Breast Parenchyma Analysis.}
%%
The analysis was conducted by calculating the cube $Parenchyma\_Percentage \sim X + Y + Z$ [\textbf{1.AF}].
%%
Using the 3D cube, \emph{KH} observed several hotspots [\textbf{2.SI}].
%%
Navigating to them using the slicing facility of the 3D visualization [\textbf{3.ZF}] highlighted features of high influence, such as image-derived features, as \emph{glandular tissue density} and \emph{parenchyma segmentation} metrics.
%%
Also, strong correlations were observed in the \emph{diabetes} slice, confirming expectations of \emph{KH} w.r.t. to its strong influence to the parenchyma tissue.
%%
A surprising find were the strong correlations with \emph{kidney disorder} ($R^2$ values around $0.9$).
%%
The [\textbf{4.DD}] analysis however showed only 8 subjects with \emph{kidney disorder}.
%%
Too few subjects impose the risk of a biased finding.
%%
The correlation was noted and will be investigated further in the near future using a extensive data set.
%%
Lastly \emph{KH} assessed the influence of contraception related features, such as use \emph{birth control pills} or the \emph{spiral}, but found no significant correlations with the parenchyma tissue.
%%
Other consumption behavior features, such as \emph{alcohol intake} also yield no elevated $R^2$ values.
%%
\emph{KH} remarked that these features are believed to have an impact on the parenchyma tissue, but they are less reliability since they are self-reported.
%%
%mamma anamnesis breast fed, contraception (spirale, pille), alcohol intake
%Lastly, \emph{KH} assessed the influence of image-derived features on the parenchyma tissue.
%%
%She observed expected results 
%Using the slicing mode to fast 
%Hypotheses
%Detailed analysis: Cholesterin, influence of image-derived feautres, classification of the BI-Rads (how well did the manual segmentation perform)
\subsection{Further Feedback and Lessons Learned} \label{Lessons Learned}
%%
Our method was well received among the domain experts.
%%
For the first time, they were able to derive an overview visualization custom tailored to underlying assumptions.
%%
\emph{KH} noted the ease of use, which "\emph{converts data sets into a feasible form}".
%%
She highlighted the efficiency of combining fast target feature selection with visually highlighting interesting results, enabling rapid analysis cycles.
%%
To get nearly similar results she would had to spend hours using \texttt{SPSS} and also might have missed many interesting hot spots during this process.
%She noted that she usually would have to spend hours using \texttt{SPSS} to assess the data in the detail we presented.
%%
\emph{TI} highlighted the ability to simultaneously analyze thousands of regression models while maintaining little time expenses for rating them.
%The novel way of modeling knowledge to derive hypothesis driven overview visualizations allow for detailed analysis of target features, either disease indicators or even image-derived features.
%%
%TI commented on the
